Synthesizing Symmetric Lenses

### 2018-07-09 papers/SymmetricLenses/main.tex
add some changes to main

In contrast to the identity lens, the \Disconnect lens does not propagate edits
from one format to the other. The disconnect lens takes four arguments: two
regular expressions and two strings. The regular expressions specify the formats
of the corresponding data on the two sides while the strings provide default
values. On creates, the input values are thrown away, and default values are
returned ($\CreateROf{\String} =
\lstinline{"UNK"}$), and on puts, the second argument is used, and the first is
thrown away ($\PutROf{\String}{\StringAlt} = \StringAlt$). In our extended
example, the \lstinline{salary} and \lstinline{company} fields are independent,
so we use the disconnect lens to relate them\bcp{Potentially confusing to
  readers (and, even to me, hacky). Why not relate each of them to the constant
  empty string?}\afm{I'm fine with the empty string.  I've seen sentinel values
  like UNK used before in practice, but if it makes the example cleaner and more
  comprehensible, sure.  Anybody else have thoughts on the matter?}\bcp{No,
  I meant why not concatenate two disconnect lenses, one relating salary to
  the empty string and one relating the empty string to company.}.
The defaults are strings that indicate the information is missing.
%
\begin{lstlisting}
disconnect(salary, company, "unk", "UNK") : salary $\Leftrightarrow$ company
\end{lstlisting}
%
The insert \lstinline{ins} and delete \lstinline{del} lenses are
syntactic sugar for uses of the disconnect lens in which a string
constant is omitted entirely from the source and target formats,
respectively.
\begin{lstlisting}
 ins($t$) = disconnect("", $t$, "", $t$)
 del($s$) = disconnect($s$, "", $s$, "")
\end{lstlisting}

@@

\begin{figure}
\begin{lstlisting}
let name_lens = id(name) . ins(",") . id(name) . del(": ") . ins(",") in
let employee_lens = name_lens . disconnect(salary,company,"unk","UNK") in
let employees_lens = ins("\n") . employee_lens . iterate(id("\n") . employee_lens) in
ins(header) . employees_lens : emp_salaries $\Leftrightarrow$ emp_insurance
\end{lstlisting}
%\ksf{I don't think this is necesssary.}
%let emp_lens = or(id(""),employees_lens) in
  \caption{A lens, written in our language for building simple symmetric lenses
    on string data, that synchronizes management and HR employee files.
    Note that we can annotate our lenses with types, to confirm the lens
    operates on the languages of the regular expressions. In this work, we
    synthesize the lens itself from such regular expressions.}
  \label{fig:example_lens}
\end{figure}

@@

We now provide the typing derivations and semantics for these lenses. For
simplicity of presentation,\bcp{Is there any known issue with relaxing this
  restriction? I've often found ti convenient to use id lenses with ambiguous
  REs...}\afm{no known issues, I just didn't want to go into specifics about
  ambiguity in an already-full paper}\bcp{OK, we can just comment that we
  believe relaxing this restriction is unproblematic.}, we define the typing derivations and
semantics only over unambiguous regular expressions, even though some lenses can
be defined over ambiguous regular expressions (for example, the identity lens
doesn't need any ambiguity constraints).

@@

\paragraph*{Relation with Simple Symmetric Lenses} To formalize how simple
symmetric lenses work as synchronization tools, we define an $apply$
function on them, much like one defines an $apply$ on classical symmetric lenses:
Given a simple symmetric lens and an optional pair of 
source and target strings,
$apply$ is a function from sequences of put objects to sequences of
put objects. If $apply(\Lens,\None,es) = es'$, then starting with no
prior data, after edit $es_i$, the lens \Lens synchronizes the other data\dpw{Is the ``other data'' $es$? or is it something else? If it is $es$ then we can say ($es$) in the sentence.} to
$es_i'$. \dpw{I wasn't sure what a ``put object'' was.  Is it a function? (If so, perhaps
call it a function and give its type?)} If $apply(\Lens,\SomeOf{(x,y)},es) = es'$, then that means, starting with
data $x$ and $y$ on the left and right, respectively, after edit $es_i$, the
lens \Lens synchronizes the other data to $es_i'$.

@@

\bcp{I guess we could also show that equivalence of simple symmetric lenses
  is just the identity relation?}
\sam{No; for example the identity lens of a concatenation of two regular expressions is the same as concatenating the two identity lenses on the respective regular expressions}

>>>>

In contrast to the identity lens, the \Disconnect lens does not propagate edits
from one format to the other. The disconnect lens takes four arguments: two
regular expressions and two strings. The regular expressions specify the formats
of the corresponding data on the two sides while the strings provide default
values. On creates, the input values are thrown away, and default values are
returned ($\CreateROf{\String} =
\lstinline{"unk"}$), and on puts, the second argument is used, and the first is
thrown away ($\PutROf{\String}{\StringAlt} = \StringAlt$). In our extended
example, the \lstinline{salary} field is only present in management files,
so we use the disconnect lens to ensure salary edits do not cause updates to the
HR file.
The defaults are strings that indicate the information is missing.
%
\begin{lstlisting}
disconnect(salary, "", "unk", "") : salary $\Leftrightarrow$ company
\end{lstlisting}
%
The insert \lstinline{ins} and delete \lstinline{del} lenses are
syntactic sugar for uses of the disconnect lens in which a string
constant is omitted entirely from the source and target formats,
respectively.
\begin{lstlisting}
 ins($t$) = disconnect("", $t$, "", $t$)
 del($s$) = disconnect($s$, "", $s$, "")
\end{lstlisting}

@@

\begin{figure}
\begin{lstlisting}
let name_lens = id(name) . ins(",") . id(name) . del(": ") . ins(",") in
let employee_lens = name_lens . disconnect(salary,"","unk","") 
                                                    . disconnect("",company,"","UNK") in
let employees_lens = ins("\n") . employee_lens . iterate(id("\n") . employee_lens) in
ins(header) . employees_lens : emp_salaries $\Leftrightarrow$ emp_insurance
\end{lstlisting}
%\ksf{I don't think this is necesssary.}
%let emp_lens = or(id(""),employees_lens) in
  \caption{A lens, written in our language for building simple symmetric lenses
    on string data, that synchronizes management and HR employee files.
    Note that we can annotate our lenses with types, to confirm the lens
    operates on the languages of the regular expressions. In this work, we
    synthesize the lens itself from such regular expressions.}
  \label{fig:example_lens}
\end{figure}

@@

We now provide the typing derivations and semantics for these lenses. For
simplicity of presentation, we define the typing derivations and semantics only
over unambiguous regular expressions, even though some lenses can be defined
over ambiguous regular expressions (for example, the identity lens doesn't need
any ambiguity constraints). However, we believe relaxing this restriction is
unproblematic.

@@

\paragraph*{Relation with Simple Symmetric Lenses} To formalize how simple
symmetric lenses work as synchronization tools, we define an $apply$ function on
them, much like one defines an $apply$ on classical symmetric lenses: Given a
simple symmetric lens and an optional pair of source and target strings, $apply$
is a function from sequences of edits to sequences of edits. If
$apply(\Lens,\None,es) = es'$, then starting with no prior data, after edit
$es_i$, the lens \Lens synchronizes the updates the data in the non-edited
format (the right format if $es_i = \InLOf{x}$, and the left format if $es_i =
\InROf{y}$) to $es_i'$. If $apply(\Lens,\SomeOf{(x,y)},es) = es'$, then that
means, starting with data $x$ and $y$ on the left and right, respectively, after
edit $es_i$, the lens \Lens synchronizes the other data to $es_i'$.

@@

\bcp{I guess we could also show that equivalence of simple symmetric lenses
  is just the identity relation?}
\sam{No; for example the identity lens of a concatenation of two regular
  expressions is the same as concatenating the two identity lenses on the
  respective regular expressions}
\afm{I think Benjamin likely means identity relation on the semantic meaning of
  the lenses, which is true.  Would this be a useful thing to prove? It's a 2
  liner.}
